---
title: "BE7023 Homework 1"
author: "Mike Lape"
date: "September 5, 2018"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---
&nbsp;

```{r, get_dat, warning=FALSE, message=FALSE}
library(faraway)
data("prostate")
```

1.	Describe the data, include size, top ten rows, and summary statistics.
```{r, fig.width=8, fig.height=8}
# The prostate dataset contains records for 97 men with prostate cancer who 
# were going to have a radical prostatectomy. 
dim(prostate)
# The data has 97 rows/observations and 9 columns/variables
# Below are the top 10 rows
head(prostate, 10)
# and here are the summary statistics for the prostate dataset.
summary(prostate)
```

&nbsp;

2.   Plot data with x = lcavol and y = lpsa.
```{r, fig.width=8, fig.height=8}
plot(prostate$lcavol, prostate$lpsa, xlab = "log(Cancer volume)", ylab = "Log(PSA)", 
     main = "Relationship between lcavol and lpsa", pch = 16 )
# The plot ranges from a little below -1 and almost up to 4 in the x-axis, and
# ranges from just below 0 to just above 5 in the y-axis.  It looks like these
# two features, lcavol and lpsa, have a pretty linear relationship with each other.
```

&nbsp;

3.  Fit simple linear regression model with y = lpsa and x = lcavol.
    Write the prediction equation.
    Report R2 and comment on it.
    Estimate population standard deviation
```{r, fig.width=8, fig.height=8}
mod <- lm(lpsa ~ lcavol, prostate)

# Get coefficients to write equation:
mod$coefficients
# Prediction Model: Yhat = 1.507 + 0.719 * X
summary(mod)$adj.r.squared
# The R2 value is 0.534, which suggests that this linear model doesn't fit the
# data very well.

# To get a good estimate of the population standard deviation we can calculate
# RMSE
sqrt(((summary(mod)$sigma)^2)/(mod$df.residual))
```

&nbsp;

4.  Prostate specific antigent (PSA) is an enzyme excreted from epithelial cells on
    the prostate.  In men with normal prostates PSA is found in the blood in small 
    quantities, but is often at a higher level in men with prostate cancer or other
    prostate issues.  It is therefore used as a diagnostic test for prostate cancer.
    By taking some blood from the man and measuring the PSA level they can determine
    if he has a helthy prostate or an abnormal one that requires further investigation.
&nbsp;

5.  Transform regression model back to original variables, comment on resultant model.
```{r}

```

&nbsp;

6.  Build 95% confidence bands as well as prediction bands around regression line.
```{r, fig.width=8, fig.height=8}
# First we need to generate some simulated lcavol lpsa pairs.
# Using min and max values to define range
sim_lcavol <-  seq(-1.3471,3.8210,0.1)
conf <- predict(mod, list(lcavol = sim_lcavol), int = "c")
pred <- predict(mod, list(lcavol = sim_lcavol), int = "p")

# plot confidence intervals
plot(prostate$lcavol, prostate$lpsa, xlab = "log(Cancer volume)", ylab = "Log(PSA)", lwd = 1)
title(main = "Scatter plot, Linear regression, 95% Confidence Bands, and 95% Prediction Bands")
matlines(sim_lcavol, conf, lty = c(1,2,2), col = c("blue","orange", "orange"), lwd = 4)
matlines(sim_lcavol, pred, lty = c(1,5,5), col = c("blue", "red", "red"), lwd = 4)
legend("topleft", legend = c("Linear Regression", "95% Confidence Band", "95% Prediction Band"),
       pch = c(16,16,16), col = c("blue", "orange", "red"))

```


